|

### 2022-05-17 Memo/memo.Rmd
edit memo

In response to comments from Reviewer 1, we have revised our Methods section to
clarify our selection of independent variables, including the use of covariates
to improve the efficiency and consistency of our estimates. Additionally, we
have included a new section in the Supplemental Materials to explain how we
conducted our survey and how it contributes to our results. We believe these
additions also give the paper a better theoretical frame to help readers sort
through the results we present.

@@

We thank Reviewer 1 for these two relevant suggestions. We now cite Gilligan and
Matsusaka's 2001 article three times in our paper. We mention it twice in our
Introduction, first on page 2: ``\textit{On the one hand, enlarged legislatures
may reduce political particularism and increase overall government spending,
which can lead to more public goods  \citep{gerring2020population,
GilliganMatsusaka2001, Weingastetal1981}}'', and then in the end of page 4:
``\textit{The corollary of the "law of $1/n$" is that increasing legislature
size promotes excessive public spending, regardless of the partisan makeup of
the legislature \citep{GilliganMatsusaka2001}}''. We also cite it in our Theory
section (page 7): ``\textit{Yet, \citet{GilliganMatsusaka2001} find that in the
US legislature size is associated primarily with higher spending on
geographically concentrated benefits, such as education and highways}''.

We cite Gerring and Veenendaal's book four times. We mention it twice on page 2:
``\textit{Legislature size is known to have a significant effect on collective
action strategies, the bargaining power between state branches, government
spending, and representation \citep{crain1979cost, freire2021legislature,
gerring2020population, PrimoSnyder2008, rogers2002free, Weingastetal1981}.
However, how legislature size impacts social welfare remains less understood. On
the one hand, enlarged legislatures may reduce political particularism and
increase overall government spending, which can lead to more public goods
\citep{gerring2020population, GilliganMatsusaka2001, Weingastetal1981}}''. We
mention the book again on page 4: ``\textit{Following the seminal contribution
by \citet{Weingastetal1981}, the "law of $1/n$" posits that larger legislatures
increase the legislative pressure for providing targeted public goods to local
supporters at the expense of the entire polity \citep{freire2021legislature}. This
question has become more relevant since legislature size has increased
consistently over time \citep{gerring2020population}}''. Finally, we also cite
it on page 6: ``\textit{Larger legislatures improve representation as both the
election thresholds and the ratio of population-per-politician decrease, which
facilitate the electoral success of ethnic and political minorities
\citep{allenrepresentation, gerring2020population}}.''

@@

We thank Reviewer 1 for the opportunity to clarify the selection of these two
sets of variables for our paper. We included an explanation for our choice on
page 17: "\textit{Our focus is on the impact of legislature size on education
and health care, which are the two most important public services the municipal
government should provide to citizens. According to the Brazilian Constitution,
municipalities must spend 25\% of their budget on education and 15\% on health
care. Consequently, the Ministry of Health and the Ministry of Education provide
readily available, fine-grained data on health care and education at the
municipal level, which we leverage in this study}."

@@

We agree with Reviewer 1 that we needed to provide more information on the
online survey. Therefore, we have detailed our procedures in Section A.14 of the
Supplementary Materials (pages 22-23), and we also reference the survey on pages
17 and 22 (footnote 13) in the main article. We rewrote paragraph 2 of Section
6.2 to clarify the importance of the survey results in our paper. It reads as
follows: ``\textit{We assume that city councilors want to extract rents and to
get reelected. The rents come from continuing in the repeated game, and the
partisanship discounts that benefit. Figure 2 shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition. This
provides further evidence that mayors do offer city councilors the benefits they
demand. In this regard, granting favors works as rent transfers for councilors.
Job appointments are also convenient for the mayors because patronage in Brazil
has positive results in terms of welfare \citep{toral2019benefits}.}''

@@

We thank Reviewer 1 for the opportunity to clarify the selection of control
variables in our empirical analysis. The inclusion of controls in our model has
two purposes. The first one is to improve the consistency of the estimates. In
Section A.5 of the Appendix, we use simulation to show that the inclusion of
covariates enhances the consistency of our point estimates. The simulation
indicates improved consistency for several distributions of the running
variables and different treatment effects. Section 5 of the manuscript
demonstrates that it improves consistency in both the main estimates and the
placebo ones. Table 1 (page 20, reproduced below) shows that without controls,
we would overestimate the treatment effects (finding an increase of 1.63
legislators instead of an increase of exactly one legislator), and underestimate
the placebo treatment effects (finding a significant decrease of 0.31
legislators where there was supposed to be no effects in legislature size). This
substantial improvement is because having control variables makes the
subpopulations in each threshold more comparable across all the thresholds,
avoiding comparing cases to the left (right) of one given threshold with points
in the right (left) of a different threshold.

@@

The second reason is that adding control variables improves the efficiency of
our estimates \citep{calonico2019regression}. We add five controls to the main
specifications. We include the 2003 population because it conditions the
multiple threshold assignments. We add GDP per capita to control for Brazil's
economic expansion during the 2000s. We also add the number of seats in 2000 to
ensure that our results come from the current legislature size and not from past
seat allocations. Year was included in the models to control for seasonal
effects. Finally, we add a dummy for Northeast states. The Northeast region
received considerable attention from the Federal Government in the 2000s.
According to \citet{de2015politica}, in 2003 the Lula administration recreated
the Superintendency for the Development of the Northeast (SUDENE), investing BRL
84 Billion from 2003 to 2010. This amounts to around 70\% of all targeted
regional investments. One example of these investments is that the Federal
Government created about 100 new universities in the Northeast region. Moreover,
infant mortality in the region decreased from 22.4 deaths per 1000 infants born
in 2005 to 19.6 deaths per 1000 infants born in 2008 due to larger federal
investments. In turn, education quality improved from 2.69 in 2005 to a 3.14
score in 2007 (0 to 10 quality of education score). For this reason, adding a
dummy for Northeast states is essential for ensuring the consistency and the
efficiency of our estimates.

To reflect these pieces of information, we rewrote paragraph five in Section 5
of the manuscript to (page 19):

``[...] \textit{we add five controls to the estimates: population in 2003; GDP
per capita; the number of seats in 2000; year; and a dummy for Northeast
municipalities. The reason for the first variable is the multiple threshold
assignments. We add GDP per capita because wealthier cities tend to have better
public services. The number of seats in 2000 intends to control that some
municipalities could experience a change provoked by the previous council size,
confounding our estimates. The Northeast dummy was added to control for the fact
that the Lula administration heavily invested in the region from 2003 until
2010, improving several development indicators \citep{de2015politica}. The years
were added to improve efficiency and control for seasonal effects. Moreover,
adding controls improves efficiency in RDD \citep{calonico2019regression} and
also enhances the consistency in the multiple thresholds RDD.}''

And we added the discussions here to the Appendix at the end of Section A.5.

We want to highlight further that this decision to add controls in our case is
also because of consistency. Even though clustered standard errors may provide
the best standard errors for our model, this would not affect the consistency.
And the consistency here relies on making the subpopulations comparable, i.e.,
the subpopulation of towns between 9 and 10 legislators as comparable as the
subpopulation of towns between 10 and 11 legislators, and so on.

@@

We did not find evidence that electing an additional legislator improves
descriptive representation in the city council. For instance, we see no
significant increase in the number of female or non-white legislators, as shown
in Table 4 (page 27). Finally, we show in the table below that the mayors'
parties are not driving our results. This result is also available on our GitHub
repository. We ommitted it from the Supplementary Materials to conserve space.

@@

We entirely agree with Reviewer 3 that we need to discuss the scope conditions
of our findings in more detail. We acknowledge that the external validity of
our results might be conditional on the contextual characteristics of our
study. We have rewritten two paragraphs in the Discussion (page 29) to highligh
the implication of our results. We write: "\textit{Our findings have
implications that extend well beyond Brazil. A general form of our argument is
that it is easy for a politician to negotiate with members of her in-group
\citep{abbink2019, alt2018giving}. Specifically, several countries have
separate executive and legislative powers, such as the federal level of
presidential systems, state governments, and local city councils. As these
institutions are influenced by political allegiances, the dynamics we expose
here may be generalizable. Depending on legislative support, changing the
legislature size could increase the chance of electing politicians that are not
aligned with the executive, hence decreasing welfare. Despite the particular
effects on political support and welfare, the logic described in this paper
would remain relevant for empirically accessing the effect of enlarged
legislatures on welfare. Our findings may also help scholars understand why
other countries in Latin America and Asia can generate effective governance
despite having strong executives and large coalitions
\citep{pereira2012surprising}.}".

@@

We would like to thank Reviewer 3 for their comments and suggestions.
\end{quote}

>>>>

In response to comments from Reviewer 1, we have revised our Methods section to
clarify our selection of independent variables, particularly the use of
covariates to improve the efficiency and consistency of our estimates. Moreover,
we have also added a new section in the Supplemental Materials to explain how we
conducted our survey and how it contributes to our results. We believe these
additions also give the paper a better theoretical frame to help readers sort
through the results we present.

@@

We thank Reviewer 1 for these two relevant suggestions. We now cite Gilligan and
Matsusaka's 2001 article three times in our paper. We mention their article
twice in our Introduction, first on page 2: ``\textit{On the one hand, enlarged
legislatures may reduce political particularism and increase overall government
spending, which can lead to more public goods  \citep{gerring2020population,
GilliganMatsusaka2001, Weingastetal1981}}'', and then on page 5: ``\textit{[The]
corollary of the "law of $1/n$" is that increasing legislature size promotes
excessive public spending, regardless of the partisan makeup of the legislature
\citep{GilliganMatsusaka2001}}''. We cite the article once more in our Theory
section (page 7): ``\textit{Yet, \citet{GilliganMatsusaka2001} find that in the
US legislature size is associated primarily with higher spending on
geographically concentrated benefits, such as education and highways}''.

We now cite Gerring and Veenendaal's book four times. We mention it twice on
page 2: ``\textit{Legislature size is known to have a significant effect on
collective action strategies, the bargaining power between state branches,
government spending, and representation \citep{crain1979cost,
freire2021legislature, gerring2020population, PrimoSnyder2008, rogers2002free,
Weingastetal1981}. However, how legislature size impacts social welfare remains
less understood. On the one hand, enlarged legislatures may reduce political
particularism and increase overall government spending, which can lead to more
public goods \citep{gerring2020population, GilliganMatsusaka2001,
Weingastetal1981}}''. We mention it again on page 4: ``\textit{Following the
seminal contribution by \citet{Weingastetal1981}, the "law of $1/n$" posits that
larger legislatures increase the legislative pressure for providing targeted
public goods to local supporters at the expense of the entire polity
\citep{freire2021legislature}. This question has become more relevant since
legislature size has increased consistently over time
\citep{gerring2020population}}''. Finally, we also cite the book on page 6:
``\textit{Larger legislatures improve representation as both the election
thresholds and the ratio of population-per-politician decrease, which facilitate
the electoral success of ethnic and political minorities
\citep{allenrepresentation, gerring2020population}}.''

@@

We thank Reviewer 1 for the opportunity to clarify why we have selected these
two sets of variables. We included an explanation for our decision on page 17:
"\textit{Our focus is on the impact of legislature size on education and health
care, which are the two most important public services the municipal government
should provide to citizens. According to the Brazilian Constitution,
municipalities must spend 25\% of their budget on education and 15\% on health
care. Consequently, the Ministry of Health and the Ministry of Education provide
readily available, fine-grained data on health care and education at the
municipal level, which we leverage in this study}."

@@

We agree with Reviewer 1 that we needed to provide more information about our
online survey. Therefore, we have detailed our survey procedures in Section A.14
of the Supplementary Materials (pages 22-23), and we also briefly discuss our
survey methodology on pages 17 and 22 (footnote 13) of the main article.
Furthermore, we have rewritten paragraph 2 of Section 6.2 to clarify the
importance of the survey results in our paper. It now reads as follows:
``\textit{We assume that city councilors want to extract rents and to get
reelected. The rents come from continuing in the repeated game, and the
partisanship discounts that benefit. Figure 2 shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition. This
provides further evidence that mayors do offer city councilors the benefits they
demand. In this regard, granting favors works as rent transfers for councilors.
Job appointments are also convenient for the mayors because patronage in Brazil
has positive results in terms of welfare \citep{toral2019benefits}.}''

@@

We thank Reviewer 1 for this constructive comment. The inclusion of control
variables in our model has two purposes. Firstly, the covariates improve the
consistency of our empirical results. In Section A.5 of the Appendix, we use
simulations to show that the inclusion of covariates enhances the consistency of
our RDD estimates. The simulations indicate that the controls improve
consistency even if we assume different distributions for the running variables
and for the treatment effects. Similaly, Section 5 of the manuscript
demonstrates that the covariates improve consistency in both the main estimates
and the placebo ones. Table 1 (page 20, reproduced below) shows that without
controls, we would overestimate the treatment effects (finding an increase of
1.63 legislators instead of an increase of exactly one legislator), and
underestimate the placebo treatment effects (finding a significant decrease of
0.31 legislators where there was supposed to be no effects in legislature size).
This substantial improvement is attributed to the addition of control variables,
which make the subpopulations in each threshold more comparable across all the
thresholds. Thus, we avoid spurious comparisons between cases on the left
(right) of one threshold and those on the right (left) of a different threshold.

@@

Secondly, the addition of control variables also improves the efficiency of our
estimates \citep{calonico2019regression}. We add five controls to the main
specifications. We include the 2003 population because it conditions the
multiple threshold assignments. We add GDP per capita to control for Brazil's
economic expansion during the 2000s. We also add the number of seats in 2000 to
ensure that our results come from the current legislature size and not from past
seat allocations. Year was included in the models to control for seasonal
effects. Finally, we add a dummy for Northeast states. The Northeast region
received considerable attention from the federal government in the 2000s.
According to \citet{de2015politica}, in 2003 the Lula administration recreated
the Superintendency for the Development of the Northeast (SUDENE), investing BRL
84 Billion from 2003 to 2010. This amounts to around 70\% of all targeted
regional investments. One example of these investments is that the federal
government created about 100 new universities in the Northeast region during
that period. Moreover, infant mortality in the Brazilian Northeast decreased
from 22.4 deaths per 1000 infants born in 2005 to 19.6 deaths in 2008 largely
due to substantial federal investments in the region. In turn, education quality
improved from 2.69 in 2005 to a 3.14 score in 2007 (0 to 10 quality of education
score). For this reason, adding a dummy for Northeast states is essential for
ensuring the consistency and the efficiency of our estimates.

To reflect these pieces of information, we have rewritten paragraph five in Section 5
of the manuscript to (page 19):

``[...] \textit{we add
five controls to the estimates: Population in 2003; GDP per capita; the number
of seats in 2000; year; and a dummy for Northeast municipalities. The reason
for the inclusion of first variable is the multiple threshold assignments. We
add GDP per capita because wealthier cities tend to have better public
services. We include number of seats in 2000 because some municipalities could
have experienced a change provoked by the previous council size, thus
confounding our estimates. The Northeast dummy was added to control for the
fact that the Lula administration heavily invested in the region from 2003
until 2010, improving several development indicators \citep{de2015politica}. We
included an indicator for year to improve efficiency and control for seasonal
effects. Moreover, adding controls improves efficiency in RDD
\citep{calonico2019regression} and also enhances the consistency in the
multiple thresholds RDD.}''

We have also added a similar discussion to Section A.5 of the Supplementary
Materials.

We want to highlight further that this decision to add controls in our case is
mainly to ensure that our estimates are consistent. Even though clustered
standard errors are adequate for our purposes, they do not improve the
consistency of our estimates. In our case, the model consistency relies on
making the subpopulations comparable. That is, the subpopulation of
municipalities who have 9 and 10 legislators should be comparable with that of
of cities between 10 and 11 legislators, and so on.

@@

We did not find evidence that electing an additional legislator improves
descriptive representation in the city council. For instance, we see no
significant increase in the number of female or non-white legislators, as shown
in Table 4 (page 27). Finally, we show in the table below that the mayors'
parties are not driving our results. We have ommitted the table from the
Supplementary Materials to conserve space, but it is available on the article's
GitHub repository.

@@

We entirely agree with Reviewer 3 that we needed to discuss the scope conditions
of our findings in more detail. We acknowledge that the external validity of
our results might be conditional on the contextual characteristics of our
study. We have rewritten two paragraphs in the Discussion (page 29) to highligh
the implication of our results. We write: "\textit{Our findings have
implications that extend well beyond Brazil. A general form of our argument is
that it is easy for a politician to negotiate with members of her in-group
\citep{abbink2019, alt2018giving}. Specifically, several countries have
separate executive and legislative powers, such as the federal level of
presidential systems, state governments, and local city councils. As these
institutions are influenced by political allegiances, the dynamics we expose
here may be generalizable. Depending on legislative support, changing the
legislature size could increase the chance of electing politicians that are not
aligned with the executive, hence decreasing welfare. Despite the particular
effects on political support and welfare, the logic described in this paper
would remain relevant for empirically accessing the effect of enlarged
legislatures on welfare. Our findings may also help scholars understand why
other countries in Latin America and Asia can generate effective governance
despite having strong executives and large coalitions
\citep{pereira2012surprising}.}".

@@

We would like to thank Reviewer 3 for their helpful comments and suggestions.
\end{quote}

### 2022-05-17 article/article.Rmd
edit memo

To avoid this problem, in section A.5 of the Supplemental Materials, we show by
simulation that adding controls, especially the variables responsible for the
multiple threshold assignments (in our case, the population in 2003), improves
the consistency and efficiency of the estimates. Relying on this fact, we add
five controls to the estimates: population in 2003; GDP per capita; the number
of seats in 2000; year; and a dummy for Northeast municipalities. The
reason for the first variable is the multiple threshold assignments. We add GDP
per capita because wealthier cities tend to have better public services. The
number of seats in 2000 intends to control that some municipalities could
experience a change provoked by the previous council size, confounding our
estimates. The Northeast dummy was added to control for the fact that the Lula
administration heavily invested in the region from 2003 until 2010, improving 
several development indicators \citep{de2015politica}. The years were added 
to improve efficiency. Moreover, adding controls improves efficiency in RDD
\citep{calonico2019regression} and also enhances the consistency in the 
multiple thresholds RDD.

>>>>

To avoid this problem, in section A.5 of the Supplemental Materials, we show by
simulation that adding controls, especially the variables responsible for the
multiple threshold assignments (in our case, the population in 2003), improves
the consistency and efficiency of the estimates. Relying on this fact, we add
five controls to the estimates: Population in 2003; GDP per capita; the number
of seats in 2000; year; and a dummy for Northeast municipalities. The reason
for the inclusion of first variable is the multiple threshold assignments. We
add GDP per capita because wealthier cities tend to have better public
services. We include number of seats in 2000 because some municipalities could
have experienced a change provoked by the previous council size, thus
confounding our estimates. The Northeast dummy was added to control for the
fact that the Lula administration heavily invested in the region from 2003
until 2010, improving several development indicators \citep{de2015politica}. We
included an indicator for year to improve efficiency and control for seasonal
effects. Moreover, adding controls improves efficiency in RDD
\citep{calonico2019regression} and also enhances the consistency in the
multiple thresholds RDD.
